---
title: "Update to ARTnet Workflow"
author: "Shawnee M. Anderson"
date: "October 15, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(MASS)
opts_chunk$set(echo=FALSE,
               cache=TRUE, autodep=TRUE, cache.comments=FALSE,
               message=FALSE, warning=FALSE)
```

## Changes to ARTnet Functions

ARTnet has been updated, and will continue to be updated, to allow for greater flexibility in modeling of network structures for use in EpiModel. 

### Geographic Stratification

ARTnet 2.0 removes the `city_name` variable and inserts a `geog.cat` variable; this variable is introduced along with the variable `geog.lvl` which specifies the level of geographic resolution desired by the user:

```{r, echo = TRUE, eval = FALSE}
epistats1 <- build_epistats(geog.lvl = "city", geog.cat = "Atlanta")
```

This new input tells `build_epistats` that the geographical level of interest is the city level and that the geographical region of interest is Atlanta. There are five levels of geographical subsetting available in ARTnet: city, state, census division and geographic region. These are called by `geog.lvl = ` city, state, div, and reg respectively in the `build_epistats` function call. An individual who would like to build a network structure based on, say, Georgia, would then use:

```{r, echo = TRUE, eval = FALSE}
epistats2 <- build_epistats(geog.lvl = "state", geog.cat = "GA")
```

Similarly, if one wishes to use the other levels of geographic stratification they can do so as above:

```{r, echo = TRUE, eval = FALSE}
epistats2 <- build_epistats(geog.lvl = "division", geog.cat = "5")
epistats3 <- build_epistats(geog.lvl = "region", geog.cat = "3")
```

If no geographic stratification is specified, `geog.lvl = NULL`, and no geopgraphic stratification is done.

```{r, echo = TRUE, eval = FALSE}
epistats4 <- build_epistats(geog.lvl = NULL)
```
### Race Stratification

A further change made to ARTnet is the ability for users to choose whether or not to add race into network estimation. Previously, this was "on"" by default; however now, with the `race = ` option in build_epistats, users are able to choose between `TRUE` or `FALSE` denoting race and no-race stratification:

```{r, echo = TRUE, eval = FALSE}
#Racial stratification
epistats4 <- build_epistats(geog.lvl = "state", geog.cat = "WA", race = TRUE)

#No racial stratification
epistats5 <- build_epistats(geog.lvl = "state", geog.cat = "GA", race = FALSE)
```

Racial stratification is set to true by default.

### Age Stratification

A further addition to the ARTnet workflow is allowing for age group subsetting and arbitrary setting of age groups. Currently ARTnet uses the full range of ages within ARTnetData and age groups are set at `0, 24, 34, 44, 54, 64, 100`. This is retained as the default behavior behavior of ARTnet with new variables `age.limits` in `build_epistats` and `age.breaks` allowing for age range and group specification respectively:

* `age.limits`: specifies the range of values to subset ARTnetData by.
* `age.breaks`: explicitly specifies the age breaks for age categories. If `NULL` defaults to 5 equally spaced categories.

```{r, echo = TRUE, eval = FALSE}
#Set age limits
epistats6 <- build_epistats(geog.lvl = "state", geog.cat = "GA", race = TRUE, age.limits = c(15, 65),
                            age.breaks = c(20, 30, 40, 50, 60))

#Specify age categories: (0-20], (20, 30], . . . (60, 100]
netparams1 <- build_netparams(epistats = epistats6, smooth.main.dur = TRUE)
```

## ARTnet Workflow

Below is the updated workflow for generating network objects for use in constructing full network parameters/statistics for use in EpiModel:

```{r, echo = TRUE, eval = FALSE}
#1. Epistats: Specify geographic features, as well as race stratification and total age range
#under consideration
epistats <- build_epistats(geog.lvl = "city", geog.cat = "Atlanta", race = TRUE, age.limits = c(30, 50),
                           age.breaks = c(35, 45))

#2. Netparams: Specify age categories if needed, or let ARTnet determine age categories by number of 
#categories desired
netparams <- build_netparams(epistats = epistats, smooth.main.dur = TRUE)

#3. Netstats: Finalize network setup 
netstats <- build_netstats(epistats, netparams, expect.mort = 0.0005, network.size = 1000, edges.avg =TRUE)

#4. Initialize network using `netstats` object from previous step.
num <- netstats$demog$num
nw <- network::network.initialize(num, directed = FALSE)

attr.names <- names(netstats$attr)
attr.values <- netstats$attr
nw <- network::set.vertex.attribute(nw, attr.names, attr.values)
nw_main <- nw_casl <- nw_inst <- nw

# 5. Main Model ---------------------------------------------------------

#Formula: 
model_main <- ~edges +
  nodematch("age.grp", diff = TRUE) +
  nodefactor("age.grp", base = 1) +
  nodematch("race", diff = FALSE) + #race = TRUE; omit if FALSE
  nodefactor("race", base = 1) +
  nodefactor("deg.casl", base = 1) +
  concurrent +
  degrange(from = 3) +
  nodematch("role.class", diff = TRUE, keep = 1:2)

# Target Stats
netstats_main <- c(
  edges = netstats$main$edges,
  nodematch_age.grp = netstats$main$nodematch_age.grp,
  nodefactor_age.grp = netstats$main$nodefactor_age.grp[-1],
  nodematch_race = netstats$main$nodematch_race_diffF, #If race = FALSE, value will be NULL
  nodefactor_race = netstats$main$nodefactor_race[-1],
  nodefactor_deg.casl = netstats$main$nodefactor_deg.casl[-1],
  concurrent = netstats$main$concurrent,
  degrange = 0,
  nodematch_role.class = c(0, 0)
)
cbind(netstats_main)
netstats_main <- unname(netstats_main)

# Fit model
fit_main <- netest(nw_main,
                   formation = model_main,
                   target.stats = netstats_main,
                   coef.diss = netstats$main$diss.byage,
                   set.control.ergm = control.ergm(MCMLE.maxit = 500,
                                                   SAN.maxit = 3,
                                                   SAN.nsteps.times = 3),
                   verbose = FALSE)

# 6. Casual Model ---------------------------------------------------------

# Formula
model_casl <- ~edges +
  nodematch("age.grp", diff = TRUE) +
  nodefactor("age.grp", base = c(1,5)) +
  nodefactor("deg.main", base = 3) +
  concurrent +
  degrange(from = 4) +
  nodematch("role.class", diff = TRUE, keep = 1:2) +
  #If race = TRUE:
  nodematch("race", diff = FALSE) +
  nodefactor("race", base = 1)

# Target Stats
netstats_casl <- c(
  edges = netstats$casl$edges,
  nodematch_age.grp = netstats$casl$nodematch_age.grp,
  nodefactor_age.grp = netstats$casl$nodefactor_age.grp[-c(1,5)],
  nodefactor_deg.main = netstats$casl$nodefactor_deg.main[-3],
  concurrent = netstats$casl$concurrent,
  degrange = 0,
  nodematch_role.class = c(0, 0),
  #If race = TRUE:
  nodematch_race = netstats$casl$nodematch_race_diffF, 
  nodefactor_race = netstats$casl$nodefactor_race[-1]
)
cbind(netstats_casl)
netstats_casl <- unname(netstats_casl)

# Fit model
fit_casl <- netest(nw_casl,
                   formation = model_casl,
                   target.stats = netstats_casl,
                   coef.diss = netstats$casl$diss.byage,
                   set.control.ergm = control.ergm(MCMLE.maxit = 500,
                                                   SAN.maxit = 3,
                                                   SAN.nsteps.times = 3),
                   verbose = FALSE)

# 7. One-Off Model --------------------------------------------------------

# Formula
model_inst <- ~edges +
  nodematch("age.grp", diff = FALSE) +
  nodefactor("age.grp", base = 1) +
  nodefactor("risk.grp", base = 5) +
  nodefactor("deg.tot", base = 1) +
  nodematch("role.class", diff = TRUE, keep = 1:2) +
  #If race = TRUE
  nodematch("race", diff = FALSE) +
  nodefactor("race", base = 1) 

# Target Stats
netstats_inst <- c(
  edges = netstats$inst$edges,
  nodematch_age.grp = sum(netstats$inst$nodematch_age.grp),
  nodefactor_age.grp = netstats$inst$nodefactor_age.grp[-1],
  nodefactor_risk.grp = netstats$inst$nodefactor_risk.grp[-5],
  nodefactor_deg.tot = netstats$inst$nodefactor_deg.tot[-1],
  nodematch_role.class = c(0, 0),
  #If race = TRUE
  nodematch_race = netstats$inst$nodematch_race_diffF,
  nodefactor_race = netstats$inst$nodefactor_race[-1]
)
cbind(netstats_inst)
netstats_inst <- unname(netstats_inst)

# Fit model
fit_inst <- netest(nw_inst,
                   formation = model_inst,
                   target.stats = netstats_inst,
                   coef.diss = dissolution_coefs(~offset(edges), 1),
                   set.control.ergm = control.ergm(MCMLE.maxit = 500,
                                                   SAN.maxit = 3,
                                                   SAN.nsteps.times = 3),
                   verbose = FALSE)

# 8. Save Data ------------------------------------------------------------

out <- list(fit_main, fit_casl, fit_inst)

saveRDS(out, file = "netest.rda")

```
